---
title: "Prescribing Information System (1993 – present)"
toc: true
toc-depth: 6
toc-title: "Variable Type"
code-tools: true
format: html
embed-resources: true
---

```{r}
#| include: false
library(tabulizer) # extracts tables from PDFs into dataframes using java
library(tidyverse) 
library(pdftools)
library(glue)
library(knitr)

# download.file("https://www.isdscotland.org/health-topics/prescribing-and-medicines/Prescribing-Datamarts/docs/PIS_fields_for_researchers_v5_eDRIS%20Guidance.pdf", "pis.pdf", mode = "wb")

#extract table of contents
pis_toc <- pdf_toc("pis.pdf") 

# extract pdf tables
pis_tables <- extract_tables("pis.pdf", output = "data.frame")



#function to make r dataframes from extracted pdf tables
pis <- c(1:5, 7:11, 12, 14, 16:20) %>%
  map(~data.frame(pis_tables[[.x]]) %>%
  select(!contains("Request")) %>%  
  select(!contains("X"))  %>%
  select(!contains(".")) %>%
  select(!ends_with("er"))  %>%
  select(!ends_with("Patient")) %>%  
  mutate(Name=ifelse(Name=="", NA, Name)) %>%
  fill(Name) %>%
  group_by(Name) %>%
  summarise(across(where(is.character), ~str_flatten(.x, collapse = " "))) %>%
  filter(!is.na(Name))  
)  

# To work on the few tables that had notes in the tables and hence were extracted poorly
pis_6 <- data.frame(pis_tables[[6]]) %>%
  select(1:2) %>%
  filter(!startsWith(Name.Format, "Note"))

pis_6$Name.Format[2] <- 
  paste(pis_6$Name.Format[2],  pis_6$Name.Format[3])

pis_6 <- pis_6 %>%
 filter(!startsWith(Name.Format, "boundary")) %>%
  mutate(Name.Format=ifelse(Name.Format=="", NA, Name.Format)) %>%
  fill(Name.Format) %>%
  group_by(Name.Format) %>%
  summarise(across(where(is.character), ~str_flatten(.x, collapse = " "))) %>%
  filter(!is.na(Name.Format)) %>%
  separate(Name.Format, c("Name", "Format"), sep="(?=Text|Nine|Eleven)",
           extra = "merge", fill = "right")


pis_21 <- data.frame(pis_tables[[21]]) %>%
  select(1:2) 
pis_21$Description.Comment[7] <- 
  paste(pis_21$Description.Comment[7],  pis_21$Description.Comment[8])

pis_21$Description.Comment[8] <- 
  paste(pis_21$Description.Comment[7])

pis_21 <- pis_21 %>%
  separate(Description.Comment, c("Description", "Comment"),
           sep="(?<=Contains dose,|typed text.)") %>%
  mutate(Comment = ifelse(is.na(Comment) & Name=="", Description, Comment)) %>%
  mutate(Description = 
           if_else(Description==Comment, NA, Description, missing = Description)) %>%
  mutate(Name=ifelse(Name=="", NA, Name)) %>%
  fill(Name) %>%
  group_by(Name) %>%
  summarise(across(where(is.character), ~str_flatten(.x, collapse = " ", na.rm = TRUE)))

pis_22 <- data.frame(pis_tables[[22]]) %>%
  select(1:2) 

saveRDS(pis, "pis.RDS")

```

```{r}
#| eval: false
#| echo: false
# run prior to Knit for easy edit 
as_glue(unlist(pis_toc))
```

[Source is this PDF from Public Health Scotland](https://www.isdscotland.org/health-topics/prescribing-and-medicines/Prescribing-Datamarts/docs/PIS_fields_for_researchers_v5_eDRIS%20Guidance.pdf)



#### 1. Prescriber

This class contains information about the prescriber

##### 1.1 Prescriber Location

Details relating to the location where the prescribing took place

```{r}
#| echo: false
kable(pis[[1]])
```

##### 1.2 Prescribing Individual

Details relating to the prescribing individual

```{r}
#| echo: false
kable(pis[[2]])
```

#### 2. Dispenser

This class contains information about the dispenser

##### 2.1 Dispenser Location

Details relating to the location where the dispensing took place

```{r}
#| echo: false
kable(pis[[3]])
```

#### 3. Patient

This class contains information about the patient

##### 3.1 Patient attributes

```{r}
#| echo: false
kable(pis[[4]])
```

#### 4. Geography/Organisation/Deprivation

##### 4.1 Geography

```{r}
#| echo: false
kable(pis[[5]])
```

##### 4.2 NHS Organisation

```{r}
#| echo: false
kable(pis_6)
```

##### 4.3 Deprivation

###### 4.3.1 Scottish Index of Multiple Deprivation (SIMD)

```{r}
#| echo: false
kable(pis[[6]])
```

###### 4.3.2 The Carstairs Deprivation Index

```{r}
#| echo: false
kable(pis[[7]])
```

#### 5. Scanned

This class contains information about prescriptions processed through the paper scanning system by Practitioner Services Division (PSD). This information includes dates, prescribable item details, form type, and quantities

##### 5.1 Prescription Date

```{r}
#| echo: false
kable(pis[[8]])
```

##### 5.2 Prescribable Item

###### 5.2.1 BNF - British National Formulary Drug Codes

```{r}
#| echo: false
kable(bind_rows(pis[[9]], pis[[10]]))
```

###### 5.2.2 Prescribable Item Details

Information relating to the prescribable item

```{r}
#| echo: false
kable(pis[[11]])
```

###### 5.2.3 Indicators

Used to identify prescribable items with specific attributes



##### 5.3 Form Type

Prescription form types

```{r}
#| echo: false
kable(pis[[12]])
```



##### 5.4 Exemption Type

Exemption types associated with prescription form

##### 5.5 Prescribed/Dispensed/Paid Items

This class contains measures and flags relating to prescribed, dispensed or paid information

```{r}
#| echo: false
kable(bind_rows(pis[[13]], pis[[14]]))
```


#### 6. Prescription Summaries

Details from 1992 until March 2004. Only summaries are available - CHI was not captured in this time period so the data is not linkable.

##### 6.1 Prescription Summaries measures

```{r}
#| echo: false
kable(pis[[15]])
```



##### 6.2 Prescription Summaries Time objects

```{r}
#| echo: false
kable(bind_rows(pis[[16]], pis[[17]]))

```


#### 7. Electronic Messaging

This class contains additional information available through electronic messaging (e-messaging)

##### 7.1 e-Prescribed Items - from April 2009 onwards.

Only available where the prescription has an ePrescribed message.

```{r}
#| echo: false
kable(bind_rows(pis_21))

```

##### 7.2 e-Dispensed Items - from December 2009 onwards.

Only available where the prescription has an eDispensed message.
